---
title: "Campus Waste and Sustainability Analytics"
author: "Shakshi"
date: "05 August 2026"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(dplyr)
library(ggplot2)
library(knitr)
```

# Introduction

This project analyzes campus waste generation and sustainability using a simulated dataset. The objective is to identify waste generation patterns, evaluate recycling performance, and suggest improvements for sustainable campus management.

# Load Dataset

```{r}
data <- read.csv("Campus_Waste_Dataset.csv", stringsAsFactors = FALSE)

head(data)

str(data)

summary(data)
```

# Data Cleaning

```{r}
sum(is.na(data))

data$Building <- as.factor(data$Building)
data$Event_Type <- as.factor(data$Event_Type)
```

# Total Waste by Building

```{r}
building_summary <- data %>%
group_by(Building) %>%
summarise(Total_Waste = sum(Total_Waste_kg))

building_summary
```

```{r}
ggplot(building_summary,
aes(x = Building, y = Total_Waste)) +
geom_col() +
theme_minimal() +
theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
labs(title = "Total Waste Generated by Building",
x = "Building",
y = "Total Waste (kg)")
```

# Waste by Event Type

```{r}
event_summary <- data %>%
group_by(Event_Type) %>%
summarise(Total_Waste = sum(Total_Waste_kg))

event_summary
```

```{r}
ggplot(event_summary,
aes(x = Event_Type, y = Total_Waste, fill = Event_Type)) +
geom_col() +
theme_minimal() +
labs(title = "Waste Generated by Event Type",
x = "Event Type",
y = "Total Waste (kg)")
```

# Recycling Performance

```{r}
average_recycled <- mean(data$Recycled_kg)

average_segregation <- mean(data$Segregation_Rate)

average_recycled

average_segregation
```

# Waste Distribution

```{r}
ggplot(data,
aes(x = Total_Waste_kg)) +
geom_histogram(bins = 15) +
theme_minimal() +
labs(title = "Distribution of Total Waste",
x = "Total Waste (kg)",
y = "Frequency")
```

# Sustainability Scorecard

```{r}
scorecard <- data %>%
group_by(Building) %>%
summarise(
Average_Waste = mean(Total_Waste_kg),
Average_Recycled = mean(Recycled_kg),
Average_Segregation = mean(Segregation_Rate)
)

scorecard
```

# Conclusion

The analysis shows that waste generation differs across campus buildings and event types. The cafeteria generates the highest amount of food waste, while the library produces more paper waste. Recycling performance can be improved through better waste segregation and awareness campaigns. These findings can help campus management improve sustainability practices.