---
title: "Campus Waste and Sustainability Analytics"
author: "Shakshi"
date: "05 August 2026"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(dplyr)
library(ggplot2)
library(knitr)
```

# Introduction

This project analyzes campus waste generation and sustainability using a simulated data set. The objective is to identify waste generation patterns, evaluate recycling performance, and suggest improvements for sustainable campus management.

# Business Analytics Framework

## Type of Analytics

This project uses **Descriptive Analytics** and **Diagnostic Analytics**.

**Descriptive Analytics** is used to summarize and understand campus waste generation, recycling performance, and waste distribution.

**Diagnostic Analytics** is used to examine differences in waste generation across campus buildings and event types and to identify areas where recycling and waste segregation can be improved.

## Specific Business Question

**Which campus buildings and event types generate the most waste, and what are the current recycling and waste segregation performance levels?**

## Required Dataset(s)

The project requires campus waste management data containing the following variables:

- `Building`
- `Event_Type`
- `Total_Waste_kg`
- `Recycled_kg`
- `Segregation_Rate`

The current project uses the simulated dataset:

`Campus_Waste_Dataset.csv`

## Plan to Access the Dataset(s)

For the current project, the dataset will be accessed as a CSV file named `Campus_Waste_Dataset.csv` and loaded into R using `read.csv()`.

For a real campus implementation, the required data can be collected from campus waste collection records, recycling records, building-wise waste reports, event-wise waste records, and waste segregation records. The data can then be combined into a CSV file with the required variables and updated regularly for analysis.

# Load Dataset

```{r}
data <- read.csv("Campus_Waste_Dataset.csv", stringsAsFactors = FALSE)

head(data)

str(data)

summary(data)
```

# Data Cleaning

```{r}
sum(is.na(data))

data$Building <- as.factor(data$Building)
data$Event_Type <- as.factor(data$Event_Type)
```

# Total Waste by Building

```{r}
building_summary <- data %>%
group_by(Building) %>%
summarise(Total_Waste = sum(Total_Waste_kg))

building_summary
```

```{r}
ggplot(building_summary,
aes(x = Building, y = Total_Waste)) +
geom_col() +
theme_minimal() +
theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
labs(title = "Total Waste Generated by Building",
x = "Building",
y = "Total Waste (kg)")
```

# Waste by Event Type

```{r}
event_summary <- data %>%
group_by(Event_Type) %>%
summarise(Total_Waste = sum(Total_Waste_kg))

event_summary
```

```{r}
ggplot(event_summary,
aes(x = Event_Type, y = Total_Waste, fill = Event_Type)) +
geom_col() +
theme_minimal() +
labs(title = "Waste Generated by Event Type",
x = "Event Type",
y = "Total Waste (kg)")
```

# Recycling Performance

```{r}
average_recycled <- mean(data$Recycled_kg)

average_segregation <- mean(data$Segregation_Rate)

average_recycled

average_segregation
```

# Waste Distribution

```{r}
ggplot(data,
aes(x = Total_Waste_kg)) +
geom_histogram(bins = 15) +
theme_minimal() +
labs(title = "Distribution of Total Waste",
x = "Total Waste (kg)",
y = "Frequency")
```

# Sustainability Scorecard

```{r}
scorecard <- data %>%
group_by(Building) %>%
summarise(
Average_Waste = mean(Total_Waste_kg),
Average_Recycled = mean(Recycled_kg),
Average_Segregation = mean(Segregation_Rate)
)

scorecard
```

# Conclusion

The analysis shows that waste generation differs across campus buildings and event types. The cafeteria generates the highest total waste among the campus buildings, with 2065 kg, while seminars generate the highest total waste among the event types, with 2083 kg. The average recycled waste is 29.883 kg and the average segregation rate is 77.405%. These findings provide useful information for campus waste management and can help identify areas that require further attention to improve sustainability practices.